Online Balanced Graph Partitioning: Optimizations for Large Clusters

### 2020-05-04 paper/disc20clustering.tex
Lower bound writing improvements

\begin{proof}
	Fix any online algorithm \ALG{}.
	Assume that $k$ is odd.
	We show how to proceed with even $k$ later in this proof.
	Let $I(z)$ denote the cluster where the node $z$ was located initially.
	Throughout the construction, we maintain ground sets for each node.
	Initially all nodes are \emph{alone} (they belong to their own trivial ground sets).

@@

	We construct a ground set containing this alone node from $S_1$ to force its eviction.
	Precisely, let $x_0$  be the alone node from $S_1$ and let $y_0$ be an arbitrary node that does not reside in $S_1$.
	We construct a ground set $\{x_0, y_0\}$.
	\ALG has them split, and the adversary produces an external request.
	$S_1$ already contains a ground set of size $k-1$, hence \ALG must collocate $\{x_0, y_0\}$ in another cluster.

	We continue to construct ground sets containing the current alone node from $S_1$.
	To preserve a feasible partition of nodes into clusters after collocating $\{x_0, y_0\}$,
	\ALG must replace $x_0$ with another single node (that we name $x_1$).
	Then, we merge the ground sets of $x_1$ and an arbitrary node $y_1$ that was initially collocated with $x_1$.
	Continuing this, we observe a sequence $X := x_0, x_1, x_2\dots$ of alone nodes that visit $S_1$.
	(Note that subsequent ground sets are different than the first ground set $\{x_0, y_0\}$, whose nodes were initially in different clusters.)

	We claim that this procedure can continue as long as a perfect partition of ground sets still exists.
	%and we construct a sequence of ground sets $X := \langle \{x_0, y_0\}, \{x_1, y_2\}, \{x_2, y_2\}, \ldots \rangle$ of length $\Omega(k \cdot \ell)$.
	Consider the moment of creating a ground set $\{x_i, y_i\}$.
	If a perfect partition of ground sets still exists, then
	\ALG{} must collocate $\{x_i,y_i\}$ in a~cluster other than $S_1$.
	To this end, it replaces $x_i$ with another alone node $x_{i+1}$.
	To continue, we must guarantee the existence of an alone node that was initially collocated with $x_{i+1}$.

	We end this procedure prematurely to keep \OPT's cost low.
	The sequence must end when no perfect partition of ground sets exists ---
	however, we end it when \emph{two more} ground sets are still possible to give.
	Precisely, we repeat constructing ground sets,
	until there are only two alone nodes left that originate from the same cluster.
	We denote these two nodes by $x^*, y^*$.
	Observe that at this point there are at most $\ell+1$ alone nodes left:
	otherwise, by pigeon hole principle, there would be more pairs of alone nodes that were initially in the same cluster.

	Now we lower bound the length of the sequence $X := x_0, x_1,\dots$.

@@

	Finally, we bound the cost of \ALG.
	\ALG incurs at least one swap for each node in the sequence $X := x_0, x_1,\dots$.

	We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
	therefore $|X| \geq k \cdot \ell - k - \ell$ \maciek{elaborate. Don't use the word "exclude" unless very obvious. Just sum up the real cost.}.
	Finally, $\ALG/\OPT \geq (k \cdot \ell - k - \ell) / 2\in \Omega(k\cdot\ell)$.
\end{proof}

>>>>

\begin{proof}
	\maciek{Better name than $S_1$?}
	Fix any online algorithm \ALG{}.
	Let $I(C)$ denote the cluster where nodes of ground set $C$ are located initially.
	If $C$ contains nodes that originate from different clusters, $I(C)$ is undefined.
	Throughout the construction, we maintain ground sets for each node.
	Initially all nodes are \emph{alone} (they belong to their own trivial ground sets).

@@

	We construct a ground set containing this alone node (and some other node).
	Precisely, let $x_0$  be the alone node from $S_1$ and let $y_0$ be an arbitrary node that does not reside in $S_1$.
	We construct a ground set $\{x_0, y_0\}$.
	\ALG has them split, and the adversary produces external requests until these are collocated.
	$S_1$ already contains a ground set of size $k-1$ (that it cannot split), hence \ALG must collocate $\{x_0, y_0\}$ in another cluster.

	We continue to construct ground sets containing the current alone node from $S_1$.
	To preserve a feasible partition of nodes into clusters after collocating $\{x_0, y_0\}$,
	\ALG must replace $x_0$ with another single node (that we name $x_1$).
	Then, we merge the ground sets of $x_1$ and the largest ground set $C$ s.t. and $C \neq \{x_0,y_0\}$~$I(C) = I(x_1)$.
	Continuing this, we observe a sequence $X := x_0, x_1, x_2\dots$ of alone nodes that visit $S_1$.
	(Note that subsequent ground sets are different than the first ground set $\{x_0, y_0\}$, whose nodes were initially in different clusters.)

	We claim that this procedure (of constructing ground sets) can continue as long as a perfect partition of ground sets still exists.
	%and we construct a sequence of ground sets $X := \langle \{x_0, y_0\}, \{x_1, y_2\}, \{x_2, y_2\}, \ldots \rangle$ of length $\Omega(k \cdot \ell)$.
	Consider the moment of creating a ground set $\{x_i, y_i\}$.
	If a perfect partition of ground sets still exists, then
	\ALG{} must collocate $\{x_i,y_i\}$ in a~cluster other than $S_1$.
	To this end, it replaces $x_i$ with another alone node $x_{i+1}$.
	To continue, we must guarantee the existence of an alone node that was initially collocated with $x_{i+1}$.
	\maciek{elaborate more}

	We end this procedure prematurely (before all nodes visit $S_1$) to keep \OPT's cost low.
	The sequence must end when no perfect partition of ground sets exists ---
	however, we end it when \emph{two more} ground sets are still possible to give.
	Precisely, we repeat constructing ground sets,
	until there are only two alone nodes left that originate from the same cluster.
	We denote these two nodes by $x^*, y^*$.
	Observe that at this point there are at most $\ell+1$ alone nodes left:
	otherwise, by pigeon hole principle, there would be more pairs of alone nodes that were initially in the same cluster.

	\ALG incurs at least one swap for each node in the sequence $X := x_0, x_1,\dots$.
	We lower bound the length of the sequence $X := x_0, x_1,\dots$.
	\maciek{TODO}
	We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
	therefore $|X| \geq k \cdot \ell - k - \ell$ \maciek{elaborate on exclude. Just sum up the real cost.}.

@@

	The competitive ratio is then: $\ALG/\OPT \geq (k \cdot \ell - k - \ell) / 2\in \Omega(k\cdot\ell)$.
\end{proof}
